---
title: "What the package does"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{What the package does}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MATH4753F25SADAT)
```


# Introduction
This R package contains a collection of functions that was collected over the course of Applied Statistical Methods. This will demonstrate how each function works and will display a result for each of them.

# The first function: myncurve()

This function accepts three values: mu, sigma, and a.


mu - represents the mean of the normal distribution


sigma - represents the standard deviation


a - is a cut-off value (or x value) where the area under a curve will be shaded


The function then produces two outputs:

1) A plot of the normal distribution with the area from $-\infty$ to 'a' shaded


2) A list containing the input values and the calculated probability

This is an example of the function being used:
```{r}
MATH4753F25SADAT::myncurve(mu=0,sigma=5, a = 6)
```



# The second function: mycltu()
This function accepts four inputs: n, iter, a, and b


n - represents the sample size used in each iteration


iter - represents the number of samples generated

a and b - represents the lower and upper bounds of the Uniform distribution

The function simulates the sampling distribution of the sample mean using a Uniform distribution. For each iteration, n values are randomly generated from the Uniform (a.b) distribution and their mean is computed. This process is repeated 'iter' times.


The function produces a histogram of the sample means.


This is an example of the function being used:
```{r}
MATH4753F25SADAT::mycltu(n = 20, iter = 10000, a = 0, b = 10)
```

# The third function: mybin()
This function accepts three inputs: iter, n, and p


iter - represents the number of simulations to be performed


n - represents the number of trials in each experiment


p - represents the probability of success for each trial


This function simulates a Binomial random process by repeatedly sampling 'n' Bernoulli trials with success probability 'p'. For each iteration, the number of successes is recorded. The process is repeated 'iter' times, producing a distribution of the number of successes.


The function creates:

1) A table of proportions for each possible number of successes from 0 to 'n'


2) A bar plot showing the empirical distribution of successes across all simulations


This is an example of the function being used:

```{r}
MATH4753F25SADAT::mybin(iter = 1000, n = 10, p = 0.5)
```



# The fourth function: mymaxlik()

The function accepts three inputs: lfun, x, and param along with optional graphical parameters (...)


x - represents the observed data


param - a vector of parameter values to be tested


lfun - a user-defined log-likelihood function that evaluates the log-likelihood for each data point


The function computes the log-likelihood for each value in 'param' by applying the log-likelihood function to every obeservation in 'x' and summing the results. These summed log-likelihood values are then plotted against the parameter values to produce a log-likelihood curve


The function identifies the maximum likelihood estimate (MLE) by locating the parameter value that produces the largest log-likelihood.


The function returns a list containing:


i: the index of the maximum log-likelihood


parami: the parameter value at the maximum


yi: the log-likelihood value at the maximum


slope: an estimate of the slope near the maximum


This is an example of the function being used:

```{r}
data <- c(3,4,3,5)
MATH4753F25SADAT::mymaxlik(x=data, param=seq(0,20,length=1000),
                   lfun=function(x,p) log(dpois(x,lambda=p)),
                   xlab=expression(lambda), main="Poisson MLE Example")
```

